---
title: "Exploratory Analysis: Automobiles"
author: "Cool Beans Programming"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading libraries and data

```{r warning=FALSE, message=FALSE}
library(ISLR2)
data(Auto)
```

## Exploring the data

```{r explore_auto}
summary(Auto)
```

All variables are quantitative except the name of the automobile. 


## Key metrics

```{r auto_range}
apply(Auto[,1:7], 2, range)
```

The ranges of the variables are given above. The greatest range is among vehicle weight. Below, the mean and standard deviation of each quantitative predictor is given.

```{r auto_mean}
apply(Auto[,1:7], 2, mean)
```

```{r auto_std}
apply(Auto[,1:7], 2, sd)
```

## Removing Observations 
```{r auto_10_85}
Auto2 <-Auto[-c(10:85),]
```


```{r auto2_range}
apply(Auto2[,1:7],2,range)
```

```{r auto2_mean}
apply(Auto2[,1:7],2,mean)
```

```{r auto2_std}
apply(Auto2[,1:7],2,sd)
```

After removing observations 10 through 85, the range, mean, and standard deviation for all quantitative variables are relatively close. 

## Investigate Predictors

```{r auto_scatterplot_matrix}
pairs(Auto[,1:7])
``` 
Many of the variables appear to have some sort of relationship though not necessarily a linear one. MPG has a non-linear relationship with most of the other predictors while displacement has a linear relationship with horsepower, weight, and acceleration. 

```{r auto_scatterplots}

par(mfrow=c(2,2))

plot(Auto$displacement, Auto$horsepower, xlab="Displacement", ylab="Horsepower")
abline(lm(Auto$horsepower ~ Auto$displacement), col="red")

plot(Auto$displacement, Auto$weight, xlab="Displacement", ylab="Weight")
abline(lm(Auto$weight ~ Auto$displacement), col="red")

plot(Auto$displacement, Auto$acceleration, xlab="Displacement", ylab="Acceleration")
abline(lm(Auto$acceleration ~ Auto$displacement), col="red")


```
## Making Predictions

To predict gas millage on the basis of the other variables, there are a few variables that may be useful in predicting mpg. 

```{r mpg_graphs}
par(mfrow=c(2,3))

plot(Auto$mpg, Auto$cylinders, xlab="Mpg", ylab="Cylinders")
plot(Auto$mpg, Auto$displacement,xlab="Mpg", ylab="Displacement")
plot(Auto$mpg, Auto$horsepower,xlab="Mpg", ylab="Horsepower")
plot(Auto$mpg, Auto$weight,xlab="Mpg", ylab="Weight")
plot(Auto$mpg, Auto$acceleration,xlab="Mpg", ylab="Acceleration")
plot(Auto$mpg, Auto$year,xlab="Mpg", ylab="Year")


```
The set of plots above clearly indicate that there is a non-linear relationship between millage and weight, displacement, and horsepower. There may be some type of relationship between millage and acceleration as well. 