---
title: "Exploratory Analysis: Boston Housing"
author: "Cool Beans Programming"
date: "`r Sys.Date()`"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the data
```{r load_boston, warning=FALSE}
library(ISLR2)
data(Boston)
```

## Explore the data
```{r boston_summary}
dim(Boston)
```

There are 13 variables and 506 observations in the Boston data set. Each row represents a different neighborhood in Boston and the columns represent the predictor variables and relate to crime rate, average number of rooms, and pupil-teacher ratio by town among others. 

## Relationship to per capita crime rate
```{r boston_predictors}
pairs(Boston)
```
Some variables like rm (average number of rooms per dwelling) and lstat (lower status of the pop.) appear to have a linear relationship with the median value of owner-occupied homes (in $1000). Overall it is difficult to tell from the scatter plot matrix. 

```{r boston_cor}
cor(Boston)[,1]
```

Rad and Tax have a strong and positive correlation to per capita crime rate while Indus, Nox, Lstat, and Age have a moderate and positive correlation with Crim. Dis and Medv have a moderate and negative correlation with Crim. 


```{r boston_hist}
hist(Boston$crim, xlab="Per Capita Crime Rate by Town")
hist(Boston$tax, xlab="Full-Value Property-Tax rate per $10,000")
hist(Boston$ptratio, xlab="Pupil-Teacher Ratio by Town")
```

The census tracts do not appear to have particularly high crime rates. The data is slightly left skewed from the pupil-teacher ratio by town. 

```{r boston_CharlesRiver}
sum(Boston$chas == 1)
```

There are 35 census tracts bound the Charles river. 

```{r boston_median_pupil}
median(Boston$ptratio)
```

The median pupil-teacher ratio by town is 19.05. 

```{r boston_subset}
t(subset(Boston, Boston$medv==min(Boston$medv)))
```
Census tracts 399 and 406 had the lowest median value of owner occupied homes.

```{r boston_rooms}
sum(Boston$rm > 8)
```

Thirteen census tracts average more than 8 rooms per dwelling. 